Clustering Beat-Chroma Patterns in a Large Music Database

### 2010-03-23 Paper/Paper.tex
improved
added title and abstract from dan

% Title.
% ------
\title{Large-scale harmonic patterns clustering}

@@

\begin{document}
%
\maketitle
%
\begin{abstract}
The goal is to analyze very large collections of music. We describe
a clustering scheme of harmonic patterns that scales linearly in time
with the number of songs analyzed. These harmonic patterns are related
to the \textit{Shingles} idea, but we scale to larger sets and we do not
perform $k$-nn with them. We create a codebook and assert
its properties. In particular, we show that clustering with notions of
bars and song segments helps, and that we get codes associated more
with some artists than others.
\end{abstract}
%
\section{Introduction}\label{sec:introduction}

@@

\begin{figure}[htb]
\begin{center}
\includegraphics[width=.9\columnwidth]{encoding_per_beat}
\end{center}
\caption{\small{Encoding larger and larger patterns with a fixed size
codebook.}}
\label{fig:perbeat}
\end{figure}

@@

\begin{figure}[htb]
\begin{center}
\includegraphics[width=.9\columnwidth]{close_patterns_diff}
\end{center}
\caption{\small{First three patterns of figure \ref{fig:cluster}
($2$nd line) presented with the centroid from Figure \ref{fig:code}
($1$st line) and the difference between both ($3$rd line).
}}
\label{fig:cluster_diff}
\end{figure}

@@

\begin{figure}[htb]
\begin{center}
\includegraphics[width=.9\columnwidth]{conf_mat_per_artist}
\end{center}
\caption{\small{Confusion matrix for artist recognition task.}}
\label{fig:conf_mat}
\end{figure}

>>>>

% Title.
% ------
\title{Clustering Beat-Chroma Patterns in a Large Music Database}

@@

\begin{document}
%
\maketitle
%
\begin{abstract}
A musical style or genre implies a set of common conventions
and patterns combined and deployed in different ways to make
individual musical pieces; for instance, most would agree that
contemporary pop music is assembled from a relatively small
palette of harmonic and melodic patterns.  The purpose of this
paper is to use a database of tens of thousands of songs
in combination with a compact representation of melodic-harmonic
content, the beat-synchronous chromagram, and data-mining
tools (clustering) to attempt to explicitly catalog this palette --
at least within the limitations of the beat-chroma representation.
We use online $k$-means clustering to summarize 3.7 million
4-beat bars in a codebook of a few hundred prototypes.
By measuring how accurately such a quantized codebook
can reconstruct the original data, we can quantify the degree
of diversity (distortion as a function of codebook size) and
temporal structure (i.e. the advantage gained
by joint quantizing multiple frames) in this music.  The most
popular codewords themselves reveal the common chords
used in the music.  Finally, the quantized representation of
music can be used for music retrieval tasks such as artist
and genre classification, and identifying songs that are
similar in terms of their melodic-harmonic content.
\end{abstract}
%
\section{Introduction}\label{sec:introduction}

@@

\begin{figure}[htb]
\begin{center}
\includegraphics[width=.9\columnwidth]{encoding_per_beat}
\end{center}
\caption{\small{Encoding larger and larger patterns with a fixed size
codebook. The size of the pattern is defined by the number of beats,
the bar information was not taken into account for that experiment.
}}
\label{fig:perbeat}
\end{figure}

@@

\begin{figure}[htb]
\begin{center}
\includegraphics[width=.8\columnwidth]{close_patterns_diff}
\end{center}
\caption{\small{First three patterns of figure \ref{fig:cluster}
($2$nd line) presented with the centroid from Figure \ref{fig:code}
($1$st line) and the difference between both ($3$rd line).
}}
\label{fig:cluster_diff}
\end{figure}

@@

\begin{figure}[htb]
%\begin{center}
\includegraphics[width=.9\columnwidth]{conf_mat_per_artist}
%\end{center}
\caption{\small{Confusion matrix for artist recognition task.}}
\label{fig:conf_mat}
\end{figure}
